---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
### 1.Code for reading in the dataset and/or processing the data - start
### Download File
```{r}
if (!file.exists("activity.zip")){ 
  download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip","Factivity.zip")
}
```
### Unzip the File 
```{r}
if (!file.exists("activity.csv")){ 
  unzip("activity.zip")
}
```
### Load the File
```{r}
allData<-read.table("activity.csv",sep= ",", header=TRUE, na.strings="NA")

```
### Preview data
```{r cache=TRUE}
head(allData)
summary(allData)
```
### Subsetting - Removing Missing Values
##6.Code to describe and show a strategy for imputing missing data -- remove the NAs
```{r}
cleanData <- na.omit(allData) 
```

### Preview data
```{r}
head(cleanData)
summary(cleanData)
```

### Format Date 
```{r}
cleanDataFormated <-data.frame(weekdays(as.Date(cleanData$date)), cleanData$date, cleanData$steps, unlist( lapply(cleanData$date,function(x) as.Date(x,origin="1970-01-01")), use.names=FALSE), cleanData$interval)
cnames<-c("weekdays","date", "steps","formatedDate","interval")
colnames(cleanDataFormated)<-cnames
inSun<-cleanDataFormated[cleanDataFormated$weekdays=='Sunday',  ]
inSat<-cleanDataFormated[cleanDataFormated$weekdays=='Saturday',  ]
inWeekend<-rbind(inSun,inSat )
inMon<-cleanDataFormated[cleanDataFormated$weekdays=='Monday',  ]
inTue<-cleanDataFormated[cleanDataFormated$weekdays=='Tuesday',  ]
inWed<-cleanDataFormated[cleanDataFormated$weekdays=='Wednesday',  ]
inThur<-cleanDataFormated[cleanDataFormated$weekdays=='Thursday',  ]
inFriday<-cleanDataFormated[cleanDataFormated$weekdays=='Friday',  ]
inWeekday<-rbind(inMon,inTue,inWed,inThur,inFriday)
#cleanDataWithWeekens <--data.frame(cleanDataFormated$steps,cleanDataFormated$interval, inWeekend)
head(inWeekend)

```

### Preview data
```{r}
head(cleanDataFormated)
summary(cleanDataFormated)
```

### aggregated total steps per day vs per seege
```{r}
aggDataTotal<- aggregate(cleanDataFormated$steps, by=list(Dates=cleanDataFormated$formatedDate), FUN=sum)
aggDataTotal

aggWeekday<- aggregate(inWeekday$steps, by=list(Dates=inWeekday$formatedDate), FUN=sum)
head(aggWeekday)
aggWeekend<- aggregate(inWeekend$steps, by=list(Dates=inWeekend$formatedDate), FUN=sum)
head(inWeekend)
```

### 2.Histogram of the total number of steps taken each day
### 7.Histogram of the total number of steps taken each day after missing values are imputed
```{r}
plot(aggDataTotal$Dates,aggDataTotal$x, type="h" , ylab="Total # of Steps per Day", xlab="", col ="black")
```

### 3.Mean and median number of steps taken each day

## What is mean total number of steps taken per day?
### Answer: Average # of steps per day excluding #NA values 
```{r}
aggDataAverage<- aggregate(cleanDataFormated$steps, by=list(Dates=cleanDataFormated$formatedDate), FUN=mean)
aggDataAverage
```
### Median # of steps  
```{r}
aggDataMedian<- aggregate(cleanDataFormated$steps, by=list(Dates=cleanDataFormated$formatedDate ), FUN=median )
aggDataMedian
```

###  4.Time series plot of the average number of steps taken -lok 
```{r}
##png(file="Average # of Stpes per Day.png",width=480,height=480)
plot(aggDataAverage$Dates,aggDataAverage$x, type="l", ylab="Average # of Stpes per Day", xlab="", col ="black")
#dev.off()


```
### 5.The 5-minute interval that, on average, contains the maximum number of steps

### aggregate # of steps by interval and date
### get the max of that 

``` {r}
aggDataTotalperInterval<- aggregate(cleanDataFormated$steps, by=list(Intervals=cleanDataFormated$interval), FUN=sum)
head(aggDataTotalperInterval)
MaxIntervalValue<- max(aggDataTotalperInterval)
head(MaxIntervalValue) 
filterTable<-data.frame(aggDataTotalperInterval)
MaxINtervalLabel<-filterTable[filterTable$x==MaxIntervalValue,1]
MaxINtervalLabel
```

So `MaxINtervalLabel` is the interval with maximum number of steps 



### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
``` {r}
plot(aggWeekday$Dates,aggWeekday$x, type="l" ,ylab="Average # of Stpes per Day Weekdays ", xlab="", col ="black")
plot(aggWeekend$Dates,aggWeekend$x, type="l",ylab="Average # of Stpes per Day Weekends", xlab="", col ="black")
```

## Are there differences in activity patterns between weekdays and weekends?

Yes 





